An empirical assessment of strategies to model opponent effects in crash severity analysis

### 2020-01-25 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
Completed first full draft of paper, to send to co-authors

  \title{An empirical assessment of strategies to model opponent effects in crash
severity analysis}
    \author[McMaster University]{Antonio Paez\corref{Corresponding Author}}
   \ead{paezha@mcmaster.ca} 
    \author[Louisiana State University]{Hany Hassan}
   \ead{hassan1@lsu.edu} 
    \author[McMaster University]{Mark Ferguson}
   \ead{fergumr@mcmaster.ca} 
    \author[McMaster University]{Saiedeh Razavi}
   \ead{razavi@mcmaster.ca} 
      \address[McMaster University]{McMaster Institute for Transportation and Logistics, 1280 Main Street
West, Hamilton, Ontario, Canada L8S 4K1}
    \address[Louisiana State University]{Department of Civil and Environmental Engineering, Baton Rouge,
Louisiana, USA 70803}

@@

\hypertarget{general-considerations}{%
\subsection{General considerations}\label{general-considerations}}

@@

\hypertarget{modelling-techniques}{%
\subsection{Modelling techniques}\label{modelling-techniques}}

@@

\hypertarget{model-specification-strategies}{%
\subsection{Model specification
strategies}\label{model-specification-strategies}}

@@

\hypertarget{data-for-empirical-assessment}{%
\subsection{Data for empirical
assessment}\label{data-for-empirical-assessment}}

@@

\hypertarget{model-estimation}{%
\subsection{Model estimation}\label{model-estimation}}

@@

\hypertarget{goodness-of-fit-of-models}{%
\subsection{Goodness of fit of models}\label{goodness-of-fit-of-models}}

@@

\begin{equation}
\label{eq:relative-likelihood}
e^{(AIC_b - AIC_b)/2}
\end{equation}

@@

\hypertarget{outcome-shares-based-on-predicted-probabilities}{%
\subsection{Outcome shares based on predicted
probabilities}\label{outcome-shares-based-on-predicted-probabilities}}

@@

\hypertarget{outcome-frequency-based-on-predicted-classes}{%
\subsection{Outcome frequency based on predicted
classes}\label{outcome-frequency-based-on-predicted-classes}}

@@

The statistics used in our assessment are summarized in Table
\ref{tab:verification-statistics}, including brief descriptions of their
interpretation. The statistics evaluate the performance of models from
different perspectives. Take as examples Percent Correct (\(PC\)) is the
sum of hits and correct rejections divided by the number of cases; Bias
(\(B\)), which measures whether a class has been over- or
under-predicted; and the probability of detection (\(POD\)) is the
fraction of hits relative to the total observed for the corresponding
class. The results of calculating the battery of verification statistics
are shown in Table \ref{tab:nowcast-outcomes-assessment} for the
nowcasting exercise, and Table \ref{tab:backcast-outcomes-assessment}
for the backcasting exercise.

@@

\hypertarget{nowcasting}{%
\subsubsection{Nowcasting}\label{nowcasting}}

First, some general remarks regarding the results are in order. It is
clear that no model performs consistently well from every perspective.
Recalling Box's maxim, all models are wrong - in this case it just so
happens that the models are wrong in subtly different ways.

The model that performs best in terms of Percent Correct and is Model 2,
followed by Model 4. The worst performer from this perspective is Model
1, with a \(PC\) score several percentage points below the top models.
The second score is Percent Correct by Class (\(PC_c\)). This score is
calculated individually for each outcome class. Model 2, again, has the
best performance for outcomes No Injury and Injury, and the second best
score for Fatality. Model 4 has the best score for Fatality, and is
second best for No Injury and Injury. Model 1 (full sample) has worst
scores for No Injury and Injury whereas its ensemble version has the
worst score for Fatality. It is important to note that Percent Correct
and Percent Correct by Class are heavily influenced by the most common
category. This can be appreciated in the scores, particularly for
Fatality, which has generally high values of the score, despite the fact
that the number of hits in that class are relatively low; the high
values in this case are due to the high ocurrence of correct rejections
elsewhere in the table.

Bias, in contrast, measures for each outcome class, the proportion of
total predictions by category (e.g., hits as well as false alarms)
relative to the total observed for that class. Overall, predictions can
have low bias (values closer to 1) but still do poorly in terms of hits.
The models with the best performance in terms of Bias are Model 1 for No
Injury and Injury, and Model 3 (ensemble) for Fatality. Model 4 is the
second best performer for No Injury and Injury, and Model 4 (ensemble)
is second best performer for Fatality. Model 1 (ensemble) has the worst
bias for No Injury and Injury, whereas Model 2 has the worst bias for
Fatality.

Critical Success Index (\(CSI\)) is a verification statistic that
assumes that the number of correct non-events is inconsequential for
determining forecasting skill. For this reason, the statistic is
calculated as the proportion of hits relative to the sum of hits plus
false alarms plus misses. No model performs uniformly best from this
perspective. Model 1 has the best \(CSI\) for No Injury, Model 2
(ensemble) has the best score for Injury, and Model 4 (ensemble) the
best score for Fatality. Model 2 has the worst score for Fatality - this
indicates that Model 4 is not particularly skilled at predicting
Fatalities correctly, given the frequency with which it gives false
alarms for this class, or misses it.

The next statistic is Probability of False Detection (\(F\)). This is
the proportion of false alarms relative to the total observed no events.
This statistic measure the frequency with which the model incorrectly
predicts an event, but not when it incorrectly misses it. The

@@

\caption{\label{tab:backcast-outcomes-assessment}\label{tab:nowcast-outcomes-assessment}Assessment of out-of-sample outcomes (backcasting using 2016 dataset)}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrllllllllll}
\toprule
\multicolumn{1}{c}{Predicted} & \multicolumn{3}{c}{Observed Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-14}
Outcome & No Injury & Injury & Fatality & \makecell[l]{Percent\\Correct} & \makecell[l]{Percent\\Correct\\by Class} & Bias$^1$ & \makecell[l]{Critical\\Success\\Index$^2$} & \makecell[l]{Probability of\\False\\Detection$^3$} & \makecell[l]{Probability\\of\\Detection$^4$} & \makecell[l]{False\\Alarm\\Ratio$^5$} & \makecell[l]{Heidke\\Skill\\Score$^6$} & \makecell[l]{Peirce\\Skill\\Score$^7$} & \makecell[l]{Gerrity\\Score$^8$}\\
\midrule
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 1}}\\
\hspace{1em}No Injury & 61684 & 27447 & 184 &  & \textcolor{red}{68.53} & \textcolor{black}{\textbf{0.9221}} & \textcolor{black}{0.4955} & \textcolor{red}{0.269} & \textcolor{black}{0.6368} & \textcolor{black}{0.3094} &  &  & \\

@@

Here I plan to discuss the applicability of the modelling strategy to
advanced modelling techniques (partial proportional odds, heterogeneity,
hierarchical models, etc.)

@@

Different modelling strategies can be used to model complex
hierarchical, multievent outcomes such as the severity of injuries
following a collision. The objective of this paper was to assess the
performance of different strategies to model opponent effects in
two-vehicle crashes. In broad terms, three strategies were considered:
1) incorporating opponent-level variables in the model; 2) single-
versus multi-level model specifications; and 3) sample subsetting and
estimation of separate models for different types of individual-opponent
interactions.

The results of the empirical assessment strongly suggest that
incorporating opponent effects can greatly improve the fit and
predictive performance of a model. There is some evidence that
subsetting the sample can improve the results in some isolated
situations (e.g., when modelling the severity of crashes involving
active travellers or motorcyclists), possibly at the risk of
overfitting. In this paper we did not compare individual models in our
Ensemble approach, but we suggest that this is an avenue for future
research.

In terms of the full sample models, the evidence was not conclusive in
favor of a single-level model with opponent variables (Model 2), or a
hierarchical model with individual-opponent interactions (Model 4). On
the one hand, the \(AIC\) tended to favor the

@@

>>>>

  \title{An empirical assessment of strategies to model opponent effects in crash
severity analysis}
    \author[McMaster University]{Antonio Paez\corref{Corresponding Author}}
   \ead{paezha@mcmaster.ca} 
    \author[Louisiana State University]{Hany Hassan}
   \ead{hassan1@lsu.edu} 
    \author[McMaster University]{Mark Ferguson}
   \ead{fergumr@mcmaster.ca} 
    \author[McMaster University]{Saiedeh Razavi}
   \ead{razavi@mcmaster.ca} 
      \address[McMaster University]{McMaster Institute for Transportation and Logistics, McMaster
University, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1}
    \address[Louisiana State University]{Department of Civil and Environmental Engineering, Louisiana State
University, Baton Rouge, Louisiana, USA 70803}

@@

\hypertarget{sec:general-considerations}{%
\subsection{General considerations}\label{sec:general-considerations}}

@@

\hypertarget{sec:techniques}{%
\subsection{Modelling techniques}\label{sec:techniques}}

@@

\hypertarget{sec:strategies}{%
\subsection{Model specification strategies}\label{sec:strategies}}

@@

\hypertarget{sec:data}{%
\subsection{Data for empirical assessment}\label{sec:data}}

@@

\hypertarget{sec:model-estimation}{%
\subsection{Model estimation}\label{sec:model-estimation}}

@@

\hypertarget{sec:goodness-of-fit}{%
\subsection{Goodness of fit of models}\label{sec:goodness-of-fit}}

@@

\begin{equation}
\label{eq:relative-likelihood}
e^{(AIC_a - AIC_b)/2}
\end{equation}

@@

\hypertarget{sec:outcomes-shares}{%
\subsection{Outcome shares based on predicted
probabilities}\label{sec:outcomes-shares}}

@@

\hypertarget{outcome-frequency-based-on-predicted-classes-secoutcomes-classes}{%
\subsection{Outcome frequency based on predicted classes
\{sec:outcomes-classes\}}\label{outcome-frequency-based-on-predicted-classes-secoutcomes-classes}}

@@

The statistics used in our assessment are summarized in Table
\ref{tab:verification-statistics}, including brief descriptions of their
interpretation (more on this below). The statistics evaluate different
aspects of the performance of a model. Some are concerned with the
ability of the model to be right. For example, Percent Correct (\(PC\))
is the sum of hits and correct rejections divided by the number of
cases. Others are concerned with the ability of the model to match the
observed outcomes, including Bias (\(B\)), which measures whether a
class has been over- or under-predicted. Whereas others measure the
ability of the model to not be wrong, such as the probability of false
detection (\(F\)) and the False Alarm Ration (\(FAR\)). We discuss the
results of calculating the battery of verification statistics, first for
the nowcasting case (are shown in (Table
\ref{tab:nowcast-outcomes-assessment}) and subsequently for the
backcasting case (Table \ref{tab:backcast-outcomes-assessment}).

@@

\hypertarget{nowcasting-verification-statistics}{%
\subsubsection{Nowcasting: verification
statistics}\label{nowcasting-verification-statistics}}

At first glance, the results of the verification statistics (Table
\ref{tab:nowcast-outcomes-assessment}) make it clear that no model under
comparison is consistently a top performer from every aspect of
prediction. Recalling Box's aphorism, all models are wrong but some are
useful - in this case it just so happens that some models are more wrong
than others in subtly different ways. That said, it is noticeable that
the worst scores across the board tend to accrue to Model 1 in its full
sample and ensemble versions. On the other hand, Model 2 (full sample)
concentrates most of the best scores and second best scores of all the
models, but also some of the worst scores for Fatality. Model 4, in
contrast, has most of the second best scores and a few top scores, but
not a single worst score.

Of all the models, Model 2 (full sample) performs best in terms of
Percent Correct, followed by Model 4 (full sample). The worst performer
from this perspective is Model 1 (full sample), with a \(PC\) score
several percentage points below the top models. The second score is
Percent Correct by Class (\(PC_c\)). This score is calculated
individually for each outcome class. Model 2 (full sample), has the best
performance for outcomes No Injury and Injury, and the second best score
for Fatality. Model 4 (full sample) has the best score for Fatality, and
is second best for No Injury and Injury. Model 1 (full sample) has worst
scores for No Injury and Injury whereas its ensemble version has the
worst score for Fatality. It is important to note that \(PC\) and
\(PC_c\) are heavily influenced by the most common category, something
that can be particularly appreciated in the scores for Fatality. The
scores for this class of outcome are generally high, despite the fact
that the number of hits are relatively low; the high values of \(PC_c\)
in this case are due to the high ocurrence of correct rejections
elsewhere in the table.

Bias (\(B\)) measures for each outcome class the proportion of total
predictions by class (e.g., hits as well as false alarms) relative to
the total number of cases observed for that class. For this reason, it
is possible for predictions to have low bias (values closer to 1) but
still do poorly in terms of hits. The models with the best performance
in terms of \(B\) are Model 1 (full sample) for No Injury and Injury,
and Model 3 (ensemble) for Fatality. Model 4 (full sample) is the second
best performer for No Injury and Injury, and Model 4 (ensemble) is
second best performer for Fatality. Model 1 (ensemble) has the worst
bias for No Injury and Injury, whereas Model 2 (full sample) has the
worst bias for Fatality.

Critical Success Index (\(CSI\)) evaluates forecasting skill while
assuming that the number of correct non-events is inconsequential skill.
Accordingly, the statistic is calculated as the proportion of hits
relative to the sum of hits plus false alarms plus misses. No model
performs uniformly best from this perspective. Model 2 (full sample) has
the best \(CSI\) for No Injury, Model 2 (ensemble) has the best score
for Injury, and Model 4 (ensemble) the best score for Fatality. On the
other hand, Model 1 (ensemble) has the worst score for No Injury, Model
1 (full sample) the worst score for Injury, and Model 2 the worst score
for Fatality. These scores indicates that the models are not
particularly skilled at predicting the corresponding classes correctly,
given the frequency with which they give false alarms or miss the class.

The next statistic is Probability of False Detection (\(F\)). This is
the proportion of false alarms relative to the total number of cases
that the event is not observed. This statistic measure the frequency
with which the model incorrectly predicts an event, but not when it
incorrectly misses it. The lowest probability of false detection in the
case of No Injury is 19.17\% for Model 2 (ensemble), with every other
model having values lower than 21\%, with the exception of Model 1 (full
sample) that has a score of 26.46\%. With respect to Injury, the lowest
probabilities range 35.8\% and 35.29\% and 35.35\% for Models 4 (full
sample) and 2 (full sample). In contrast, the highest probability of
false detection for Injury is 45.12\% for Model 1 (ensemble). The scores
for \(F\) for Fatality are all extremely low as a consequence of the
very low frequency of this class of outcome in the sample.

The Probability of Detection (\(POD\)) measures the frequency with which
the model correctly predicts a class, relative to the number of cases of
that class. As with some other verification statistics, no model is
consistently a best performer. Model 4 (full sample) has the highest
probability of detection for No Injury (65.38\%), followed by Model 2
(65.32\%), whereas the worst probability of detection is by Model 1
(ensemble) with a score of 55.54\%. In terms of Injury, all models have
\(POD\) higher than 79\%, and the highest score is 80.67\% for Model 2
(ensemble). The exception is Model 1 (full sample), which has a
considerably lower \(POD\) of Injury with a score of 73.36\%. Lastly, in
terms of Fatalities, all models have very low probabilities of
detection, ranging from a high of 4.94\% in the case of Model 4
(ensemble) to a worst score of 0.11\% in the case of Model 2 (full
sample).

The False Alarm Ratio (\(FAR\)) is the fraction of predictions by class
that were false alarms, which means that lower scores are better. Model
2 (full sample) has the best \(FAR\) statistic for No Injury, as only
25.05\% of predictions for this class are false alarms. The next best
score is by Model 4 (full sample), with only 25.23\% of No Injury
predictions being false alarms. The worst performance in this class is
by Model 1 (ensemble), which produces almost a third of false alarms in
its predictions of No Injury. In the case of Injury, the False Alarm
Ratio ranges from a low of 29.15\% by Model (ensemble), with every other
model having scores lower than 30\% except Model 1 (full sample), that
gives almost 32\% of false alarms. In terms of Fatality, the lowest
\(FAR\) is also for Model 4 (ensemble) with only 17.86\% of false
alarms, whereas the worst performance is by Model 2 (full sample), which
produces over 95\% of false alarms.

The next three verification statistics are skill scores that
simultaneously consider different aspects of prediction, and are
therefore overall indicators of prediction skill. Heidke's Skill Score
(\(HSS\)) is the fraction of correct predictions above those that could
be attributed to chance. Peirce's Skill Score (\(PSS\)) combines the
Probability of Detection (\(POD\)) of a model and its Probability of
False Detection (\(F\)) to measure the skill of a model to discriminate
the classes of outcomes. Lastly, Gerrity Score (\(GS\)) is a measure of
the model's skill predicting the correct classes that tends to reward
correct forecasts of the least frequent class. These scores help to
remove some of the ambiguity regarding the overall performance of a
model. In this way, we know that Model 2 (full sample) does not do
particularly well with the class Fatality - however, of all models, it
tends to have the best overall performance. Its \(HSS\), for example,
suggests that it achieves 44.74\% of correct predictions after removing
correct predictions attributable to chance. In contrast, the lowest
score is for Model 1 (ensemble), which only achieves 36.26\% correct
predictions after removing those attributable to chance. Model 2 (full
sample) also has the highest \(PSS\) and the second highest \(GS\).
Model 4 (full sample) has the highest \(GS\) and the second highest
\(HSS\) and \(PSS\).

@@

\hypertarget{backcasting-verification-statistics}{%
\subsubsection{Backcasting: verification
statistics}\label{backcasting-verification-statistics}}

Table \ref{tab:backcast-outcomes-assessment} presents the results of the
verification exercise for the case of our out-of-sample predictions
(i.e., backasting). Qualitatively, the results are similar to those of
the nowcasting experiments, but with a somewhat weaker performance of
the ensemble models. This, again, supports the idea that these models
might be overfitting the process, as discussed in reference to the
aggregate forecasts (see Section \ref{sec:outcomes-shares}). Models 2
(full sample) and 4 (full sample) are again identified as the best
overall performers, and particularly Model 2 (full sample) performs
somewhat more adroitly with respect to Fatality in backcasting than it
did in nowcasting.

\caption{\label{tab:backcast-outcomes-assessment}\label{tab:backcast-outcomes-assessment}Assessment of out-of-sample outcomes (backcasting using 2016 dataset)}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrllllllllll}
\toprule
\multicolumn{1}{c}{Predicted} & \multicolumn{3}{c}{Observed Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-14}
Outcome & No Injury & Injury & Fatality & \makecell[l]{Percent\\Correct} & \makecell[l]{Percent\\Correct\\by Class} & Bias$^1$ & \makecell[l]{Critical\\Success\\Index$^2$} & \makecell[l]{Probability of\\False\\Detection$^3$} & \makecell[l]{Probability\\of\\Detection$^4$} & \makecell[l]{False\\Alarm\\Ratio$^5$} & \makecell[l]{Heidke\\Skill\\Score$^6$} & \makecell[l]{Peirce\\Skill\\Score$^7$} & \makecell[l]{Gerrity\\Score$^8$}\\
\midrule
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 1}}\\
\hspace{1em}No Injury & 61684 & 27447 & 184 &  & \textcolor{red}{68.53} & \textcolor{black}{\textbf{0.9221}} & \textcolor{black}{0.4955} & \textcolor{red}{0.269} & \textcolor{black}{0.6368} & \textcolor{black}{0.3094} &  &  & \\

@@

As discussed in Section \ref{sec:techniques}, there is a rich selection
of modelling approaches that are applicable to crash severity analysis.
Based on the literature, we limited our empirical assessment of
modelling strategies to only one model, namely the ordinal probit. On
the other hand, since the modelling strategies discussed here all relate
to the specification of the latent function and data subsetting, it is a
relatively simple matter to extend them to other modelling approaches.
For example, take Expression
\ref{eq:latent-function-with-opponent-variables} and add a random
component \(\mu_{k}\) as follows:

\begin{equation}
\label{eq:latent-function-with-opponent-variables-and-random-component}
y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \sum_{r=1}^R\delta_ro_{jvkr} + \mu_{k} + \epsilon_{itk}
\end{equation}

The addition of the random component in this fashion would help to
capture, when appropriate, unobserved heterogeneity at the level of the
crash (this is similar to the random intercepts approach in multi-level
modelling; also see Mannering et al., 2016). As a second example, take
Expressions \ref{eq:latent-function-with-hierarchical-traffic-unit} to
\ref{eq:hierarchical-traffic-unit-coefficients} and add a random
component to a hierarchical coefficient, to obtain:

\begin{equation}
\label{eq:hierarchical-traffic-unit-coefficients-with-random-component}
\begin{array}{rcl}\
\beta_{m}u_{tkm} &=& \big( \beta_{m1} + \beta_{m2}p_{itk2} + \cdots + \beta_{mL}p_{itkL} + \mu_{mk}\big)u_{tkm}\\ 
&=& \beta_{m1}u_{tkm} + \beta_{m2}p_{itk2}u_{tkm} + \cdots + \beta_{mL}p_{itkL}u_{tkm} + \mu_{mk}u_{tkm}
\end{array}
\end{equation}

This is similar to the random slopes strategy in multi-level modelling.

We do not report results regarding other modelling strategies. On the
one hand, more sophisticated modelling frameworks are generally capable
of improving the performance of a model. On the other hand, there are
well-known challenges in the estimation of more sophisticated models
(see Lenguerrand et al., 2006, p. 47, for a discussion of convergence
issues in models with mixed effects; Mannering et al., 2016, p. 13, for
some considerations regarding the complexity and cost of estimating more
complex models; and Bogue et al., 2017, p. 27, on computational demands
of models with random components). The additional cost and complexity of
more sophisticated modelling approaches would, in our view, have greatly
complicated our empirical assessment, particularly considering the large
size of the sample involved in this research (a dataset with over
164,000 records in the case of the full sample models). That said, we
experimented with a model with random components using monthly subsets
of data to find that, indeed, estimation takes considerably longer, is
more demanding in terms of fixing potential estimation quirks, and in
the end resulted in variance components that could not be reliably
estimated as different from zero (results can be consulted in the source
R Notebook). For this reason, we choose to leave the application of more
sophisticated models as a matter for future research.

@@

The study of crash severity is an important component of accident
research, as seen from a large and vibrant literature and numerous
applications. Part of this literature covers different modelling
strategies that can be used to model complex hierarchical, multievent
outcomes such as the severity of injuries following a collision. In this
paper, our objective has been to assess the performance of different
strategies to model opponent effects in two-vehicle crashes. In broad
terms, three strategies were considered: 1) incorporating opponent-level
variables in the model; 2) single- versus multi-level model
specifications; and 3) sample subsetting and estimation of separate
models for different types of individual-opponent interactions. The
empirical evaluation was based on data from Canada's National Crash
Database and the application of ordered probit models. A suite of models
that implemented the various strategies considered was estimated using
data from 2017. We then assessed the performance of the models using one
information criterion (AIC). Furthermore, the predictive performance of
the models was assessed in terms of both nowcasting (in-sample
predictions) and backcasting (out-of-sample predictions), the latter
using data from 2016.

The results of the empirical assessment strongly suggest that
incorporating opponent effects can greatly improve the goodness-of-fit
and predictive performance of a model. Two modelling strategies appear
to outperform the rest: a relatively simple single-level modelling
approach that incorporates opponent effects, and a hierarchical
modelling approach with nested opponent effects. There was some evidence
that subsetting the sample can improve the results in some cases s(e.g.,
when modelling the severity of crashes involving active travellers or
motorcyclists), but possibly at the risk of overfitting the process and
leading to worse out-of-sample performance. In this paper we did not
compare individual models in our ensemble approach, but we suggest that
this is an avenue for future research.

The results of this research should be informative to analysts
interested in crashes involving two parties, since it provides some
useful guidelines regarding the specification of opponent effects.

The analysis also opens up a few avenues for future research. First, for
reasons discussed in Section \{sec:further-considerations\}, we did not
consider more sophisticated modelling approaches, such as models with
random components, partial proportional odds, ranked ordered models, or
multinomial models, to mention just a few possibilities. Secondly, we
only considered the performance of the models when making predictions
for the full sample. That is, the submodels in the ensembles were not
compared in detail, just their aggregate results when predicting the
full sample. However, the goodness-of-fit was not uniformly better for
any one modelling strategy when the data were subset, and it is possible
that individual models perform better for a certain subset than
competitors that are part of a better ensemble, overall. For this
reason, we suggest that additional work with ensemble approaches is
warranted. Finally, it is clear that the models do not generally do well
when predicting the least frequent class of outcome, namely Fatality. It
would be worthwhile to further investigate approaches for so-called
imbalanced learning, a task where Tarrao et al.~(2014) have already made
some headway.

Finally, as an aside, this paper is, to the best of our knowledge, the
first example of reproducible research in crash severity analysis. By
providing the data \emph{and} code for the analysis, it is our hope that
this will allow other researchers to easily verify the results, and to
extend them. A common practice in the machine learning community is to
use canonical dataset to demonstrate the performance of new techniques.
Sharing code and data has remained relatively rare in transportation
research, and we would like to suggest that the datasets used in this
research could constitute one such canonical dataset for future
methodological developments.

\hypertarget{acknowledgments}{%
\section*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{section}{Acknowledgments}

To be added in the final version of the paper.

@@

\leavevmode\hypertarget{ref-Lenguerrand2006modelling}{}%
Lenguerrand, E., Martin, J.L., Laumon, B., 2006. Modelling the
hierarchical structure of road crash data - application to severity
analysis. Accident Analysis and Prevention 38, 43--53.
doi:\href{https://doi.org/10.1016/j.aap.2005.06.021}{10.1016/j.aap.2005.06.021}
